RDF Validation Requirements - Evaluation and Logical Underpinning

### 2015-01-09 paper/RDF validation requirements/RDF validation requirements.tex
finalize paper

\ms{Intersection} (composition, conjunction) corresponds to the requirements \ms{R-15-CONJUNCTION-OF-CLASS-EXPRESSIONS} and \ms{R-16-CONJUNCTION-OF-DATA-RANGES}.
DLs allow new concepts and roles to be
built using a variety of different constructors. We distinguish concept and role 
constructors depending on whether concept or role expressions are constructed. In the case of
concepts, one can further separate basic Boolean constructors, role restrictions and nominals/enumerations \cite{Kroetzsch2012}.
Boolean concept constructors provide basic boolean operations that are closely related to
the familiar operations of intersection, union and complement of sets, or to conjunction,
disjunction and negation of logical expressions \cite{Kroetzsch2012}.

@@

requirement:

\begin{itemize}
	\item R-50-WHITESPACE-HANDLING-OF-RDF-LITERALS
\end{itemize}

\begin{itemize}
	\item avoid whitespaces in literals neither leading nor trailing white spaces
\end{itemize}

@@

check if literals of the property \ms{p} do not include whitespaces

@@

requirement:

\begin{itemize}
	\item R-51-HTML-HANDLING-OF-RDF-LITERALS
\end{itemize}

\begin{itemize}
	\item check if there are no HTML tags included in literals (of specific data properties within the context of specific classes)
\end{itemize}

@@

check if literals of the property \ms{p} of the class \ms{C} do not include HTML tags

@@

requirements:

\begin{itemize}
	\item R-52-NEGATIVE-OBJECT-PROPERTY-CONSTRAINTS
	\item R-53-NEGATIVE-DATA-PROPERTY-CONSTRAINTS
\end{itemize}

\begin{itemize}
	\item instances of a specific class must not have some object property
  \item OWL 2 DL: ObjectComplementOf ( ObjectSomeValuesFrom ( ObjectPropertyExpression owl:Thing ) )
\end{itemize}

@@

\begin{ex}
# ShEx
# ----
<FeelingForce> {
    :feelingForce (true) ,
    :attitute xsd:string }
<JediMentor> {
    :feelingForce (true) ,
    :attitute ('good') ,
    :laserSwordColor xsd:string ,
    :numberLaserSwords xsd:nonNegativeInteger ,
    :mentorOf @<JediStudent> ,
   !:studentOf @<JediMentor> }
<JediStudent> {
    :feelingForce (true) ,
    :attitute ('good') ,
    :laserSwordColor xsd:string ,
    :numberLaserSwords xsd:nonNegativeInteger ,
   !:mentorOf @<JediStudent> ,
    :studentOf @<JediMentor> }
\end{ex}

\begin{itemize}
	\item A matching triple has any predicate except those excluded by the '!' operator.
  \item \url{http://w3.org/brief/Mzk0}
\end{itemize}

\begin{ex}
# individuals matching 'FeelingForce' and 'JediMentor' data shapes
# ----------------------------------------------------------------
:Obi-Wan 
    :feelingForce true ;
    :attitute 'good' ;
    :laserSwordColor 'blue' ;
    :numberLaserSwords '1'^^xsd:nonNegativeInteger ;
    :mentorOf :Anakin .
\end{ex}

\begin{ex}
# individuals matching 'FeelingForce' and 'JediStudent' data shapes
# -----------------------------------------------------------------
:Anakin 
    :feelingForce true ;
    :attitute 'good' ;
    :laserSwordColor 'blue' ;
    :numberLaserSwords '1'^^xsd:nonNegativeInteger ;
    :studentOf :Obi-Wan .
\end{ex}

@@

requirement:

\begin{itemize}
	\item R-120-HANDLE-RDF-COLLECTIONS
\end{itemize}

examples:

\begin{itemize}
	\item size of collection
	\item first / last element of list must be a specific literal
	\item compare elements of collection
	\item are collections identical?
	\item actions on RDF lists: \url{http://www.snee.com/bobdc.blog/2014/04/rdf-lists-and-sparql.html}
	\item 2. list element equals ''
	\item Does the list have more than 10 elements?
\end{itemize}

@@

get 2. list element

\begin{ex}
# SPIN
# ----
:getListItem
    a spin:Function ; rdfs:subClassOf spin:Functions ;
    spin:constraint [
        rdf:type spl:Argument ;
        spl:predicate sp:arg1 ;
        spl:valueType rdf:List ;
        rdfs:comment "list" ; ] ;
    spin:constraint [
        rdf:type spl:Argument ;
        spl:predicate sp:arg2 ;
        spl:valueType xsd:nonNegativeInteger ;
        rdfs:comment "item position (starting with 0)" ; ] ;
    spin:body [
        a sp:SELECT ;
        sp:text """
            SELECT ?item
            WHERE {
                ?arg1 :contents/rdf:rest{?arg2}/rdf:first ?item } """ ; ] ;
    spin:returnType rdfs:Resource .
\end{ex}

\begin{ex}
# data
# ----
:Jinn :students 
     ( :Xanatos :Kenobi ) . 
\end{ex}

\begin{ex}
# SPIN
# ----
BIND ( :getListItem( ?list, "1"xsd:nonNegativeInteger ) AS ?listItem ) .
\end{ex}

@@

\begin{ex}
# result
# ------
:Kenobi
\end{ex}

@@

requirements:

\begin{itemize}
	\item R-222-RECURSIVE-QUERIES
\end{itemize}

\begin{itemize}
	\item If we want to define Resource Shapes, remember that it is a recursive
language (the valueShape of a Resource Shape is in turn another
Resource Shape). There is no way to express that in SPARQL without
hand-waving "and then you call the function again here" or "and then
you embed this operation here" text.  The embedding trick doesn't work
in the general case because SPARQL can't express recursive queries,
e.g. "test that this Issue is valid and all of the Issues that
references, recursively".
  \item Most SPARQL engines already have
functions that go beyond the official SPARQL 1.1 spec. The cost of that
sounds manageable to me.
\end{itemize}

@@

\begin{ex}
# ShEx
# ----
IssueShape {
    related @IssueShape*
}
\end{ex}

@@

requirements:

\begin{itemize}
	\item 
\end{itemize}

@@

\begin{itemize}
	\item a date is really a date
  \item SPARQL regex can be used
\end{itemize}

@@

\begin{ex}
# SPIN
# ----
FILTER ( datatype( ?shoeSize ) = xsd:nonNegativeInteger )
isNumeric ( ?shoeSize )
\end{ex}

\begin{itemize}
	\item datatype of ?showSize id xsd:nonNegativeInteger
	\item datatype is really numeric
\end{itemize}

@@

\begin{DL} 
$\rho(\textsf{hasBirthday}) \sqsubseteq \PVerb{xsd:dateTime}$\\
\end{DL}

Equal to the definition of a data property range.

\tb{I do not understand the relation of this example to this constraint}

@@

\begin{ex}
# OWL 2
# -----
:hasBrother owl:equivalentProperty :hasMaleSibling . 
:Chris :hasBrother :Stewie . 
:Stewie :hasMaleSibling :Chris .
\end{ex}

@@

\begin{ex}
:Chris :hasMaleSibling a:Stewie . 
:Stewie :hasBrother a:Chris .
\end{ex}

\begin{DL}
:hasBrother $\sqsubseteq$ :hasMaleSibling $\sqcap$ :hasMaleSibling $\sqsubseteq$ :hasBrother \\
\end{DL}

@@

\begin{itemize}
	\item positive property assertions
  \item negative property assertions
	\item A positive object property assertion ObjectPropertyAssertion( OPE a1 a2 ) states that the individual a1 is connected by the object property expression OPE to the individual a2. 
	\item A negative object property assertion NegativeObjectPropertyAssertion( OPE a1 a2 ) states that the individual a1 is not connected by the object property expression OPE to the individual a2. 
	\item A positive data property assertion DataPropertyAssertion( DPE a lt ) states that the individual a is connected by the data property expression DPE to the literal lt. 
	\item A negative data property assertion NegativeDataPropertyAssertion( DPE a lt ) states that the individual a is not connected by the data property expression DPE to the literal lt.
\end{itemize}

@@

\begin{ex}
# OWL 2
# -----
NegativeObjectPropertyAssertion( :hasSon :Peter :Meg )
\end{ex}

@@

\begin{ex}
# OWL 2
# -----
DataPropertyAssertion( :hasAge :Meg "17"^^xsd:integer )
\end{ex}

@@

\begin{ex}
# OWL 2
# -----
:hasFather rdf:type owl:FunctionalProperty . 	
:Stewie :hasFather :Peter . 	
:Stewie :hasFather :Peter_Griffin . 
\end{ex}

@@

\begin{ex}
:Peter_Griffin owl:sameAs :Peter .  
\end{ex}

@@

An object property inverse functionality axiom InverseFunctionalObjectProperty( OPE ) states that the object property expression OPE is inverse-functional — that is, for each individual x, there can be at most one individual y such that y is connected by OPE with x. Each such axiom can be seen as a syntactic shortcut for the following axiom: SubClassOf( owl:Thing ObjectMaxCardinality( 1 ObjectInverseOf( OPE ) ) ).

@@

\begin{ex}
# OWL 2
# -----
:fatherOf rdf:type owl:InverseFunctionalProperty . 	
:Peter :fatherOf :Stewie .
:Peter_Griffin :fatherOf :Stewie .
\end{ex}

@@

\begin{ex}
:Peter owl:sameAs :Peter_Griffin . 
\end{ex}

@@

\begin{ex}
# OWL 2
# -----
:Peter :fatherOf :Stewie . 
ObjectHasValue( :fatherOf :Stewie )
\end{ex}

The has-value class expression contains those individuals that are connected through the :fatherOf property with the individual :Stewie.
:Peter is classified as its instance.

@@

\begin{ex}
# OWL 2
# -----
:Peter :likes :Peter . 
ObjectHasSelf( :likes ) 
\end{ex}

\begin{DL}
LikesThemselves $\sqsubseteq$ $\exists$ likes . Self
\end{DL}

@@

The self-restriction contains those individuals that like themselves.
:Peter is classified as its instance.

@@

requirement:

\begin{itemize}
	\item R-46-CONSTRAINING-FACETS
\end{itemize}

\begin{itemize}
  \item For datatype properties it should be possible to declare frequently needed "facets" to drive user interfaces and validate input against simple conditions, including min/max value, regular expressions, string length etc. similar to XSD datatypes. 
	\item constraining facets to restrict datatypes of RDF literals
  \item constraining facets: xsd:length, xsd:minLength, xsd:maxLength, xsd:pattern, xsd:enumeration, xsd:whiteSpace, xsd:maxInclusive, xsd:maxExclusive, xsd:minExclusive, xsd:minInclusive, xsd:totalDigits, xsd:fractionDigits

\end{itemize}

@@

\begin{ex}
# OWL 2 QL (functional-style syntax)
# ----------------------------------
Declaration( Datatype( :SSN ) ) 
DatatypeDefinition( 
    :SSN
    DatatypeRestriction( xsd:string xsd:pattern "[0-9]{3}-[0-9]{2}-[0-9]{4}" ) )     
DataPropertyRange( :hasSSN :SSN ) 
\end{ex}

\begin{ex}
# OWL 2 QL (turtle syntax)
# ------------------------
:SSN 
    a rdfs:Datatype ;
    owl:equivalentClass [
        a rdfs:Datatype ;
        owl:onDatatype xsd:string ;
        owl:withRestrictions ( 
            [ xsd:pattern "[0-9]{3}-[0-9]{2}-[0-9]{4}" ] ) ] .
:hasSSN rdfs:range :SSN .
\end{ex}

\begin{itemize}
	\item A social security number is a string that matches the given regular expression. 
The second axiom defines :SSN as an abbreviation for a datatype restriction on xsd:string. 
The first axiom explicitly declares :SSN to be a datatype. 
The datatype :SSN can be used just like any other datatype; 
for example, it is used in the third axiom to define the range of the :hasSSN property. 
\end{itemize}

\begin{ex}
# valid data
# ----------
:TimBernersLee
    :hasSSN "123-45-6789"^^:SSN .
\end{ex}

\begin{ex}
# invalid data
# ------------
:TimBernersLee
    :hasSSN "123456789"^^:SSN .
\end{ex}

@@

\subsection{Simple Example}
Each object is uniquely identified by its social security number.

\begin{ex}
# DSP
# ---
-
\end{ex}

\begin{ex}
# OWL2 DL
# -------
owl:Thing owl:hasKey ( :hasSSN ) . 	
\end{ex}

\begin{ex}
# ReSh
# ----
-
\end{ex}

\begin{ex}
# ShEx
# ----
-
\end{ex}

The first axiom makes :hasSSN the key for instances of the owl:Thing class; thus, only one individual can have a particular value for :hasSSN.

\begin{ex}
# data
# ----
:Peter :hasSSN "123-45-6789" . 
:PeterGriffin :hasSSN "123-45-6789" . 
\end{ex}

\begin{DL}
$(\textsf{funct } \textsf{SSN}\sp{\overline{\ }})$
\end{DL}

\begin{gcotable}
property & $\top$ & SSN$^{-}$ & SSN & - & inverse & - \\
property & $\top$ & SSN$^{-}$ & - & - & functional & - \\
\end{gcotable}

Since the values of :hasSSN are the same for the individuals a:Peter and a:PeterGriffin, these two individuals are equal - that is, this ontology entails the following assertion :Peter owl:sameAs :PeterGriffin .

@@

Since "Each object is uniquely identified by its social security number." It means it is inverse functional property.

So instead,  a property can be defined as 
\begin{DL}
$(\textsf{funct } \textsf{SSN})$
\end{DL}
, and then by a property inclusion axiom

\begin{DL}
$haskey \sqsubseteq SSN^{-}$
\end{DL}

now for a given number haskey will give me the object (note that obviously it should be named, no ambigious individuals).

However I needed to do some little research and found out that, keys are even more general (A generalisation of inverse functional properties) so that it can also be a chain of properties. (it is clear here: http://www.w3.org/TR/2012/REC-owl2-rdf-based-semantics-20121211/\#table-semcond-keys)

For this generalisation purposes and many spooky interactions of different sorts of key (and with some elaboration they can lead to undecidability), they introduced "key boxes, and extended the language with special "keyfor" construct. So how it is done in DL community as I learned is

\begin{DL}
SSN "keyfor" Human $\sqcap$ TurkishCitizen
\end{DL}

 for instance. Now ssn behaves as an identifying element for individuals of right hand side.

For comprehensive studies and resources, one can find lots of examples in:

http://lat.inf.tu-dresden.de/research/mas/Din-Mas-06.pdf
https://www.jair.org/media/1542/live-1542-2375-jair.pdf

Yes you can, but as you see it is not restriction statement by its own. Nevertheless, I think, in DL mapping, use of "keyfor" and sending a reference to one of those papers is more convenient/safer, since it can be a datatype property, object property, or a chain of properties (from OWL world), as well as DL literature has it as "keyfor".

\begin{gcotable}
property & $\top$ & SSN & - & - & keyfor & - \\
\end{gcotable}

@@

requirement:

\begin{itemize}
	\item R-224-instance-level-data-validation-exploitingclass/property-specialization-axioms-in-ontologies
\end{itemize}

\begin{itemize}
  \item Validation of instances data (direct or indirect) exploiting the subclass or sub-property link in a given ontology.
  \item This validation can indicate when the data is verbose (redundant) or expressed at a too general level, and could be improved.
\end{itemize}

examples:

\begin{itemize}
	\item If dc:date and one of its sub-properties dcterms:created or dcterms:issued are present, check that the value in dc:date is not redundant with dcterms:created or dcterms:issued for ingestion
  \item Check if dc:rights has the same value than edm:rights either as rdf:resource or literal, if yes dc:rights is redundant
  \item If one or more dc:coverage are present, suggest the use of one of its sub-properties, dcterms spatial or dcterms:temporal.
\end{itemize}

>>>>

\ms{Intersection} (composition, conjunction) corresponds to the requirements \\
\ms{R-15-CONJUNCTION-OF-CLASS-EXPRESSIONS} and \ms{R-16-CONJUNCTION-OF-DATA-} \ms{RANGES}.
DLs allow new concepts and roles to be
built using a variety of different constructors. We distinguish concept and role 
constructors depending on whether concept or role expressions are constructed. In the case of
concepts, one can further separate basic Boolean constructors, role restrictions and nominals/enumerations \cite{Kroetzsch2012}.
Boolean concept constructors provide basic boolean operations that are closely related to
the familiar operations of intersection, union and complement of sets, or to conjunction,
disjunction and negation of logical expressions \cite{Kroetzsch2012}.

@@

\ms{Whitespace Handling} corresponds to the requirement
\ms{R-50-WHITESPACE-} \\
\ms{HANDLING-OF-RDF-LITERALS}.
Avoid whitespaces in literals neither leading nor trailing white spaces.

@@

Check if literals of the property \ms{p} do not include whitespaces.

@@

\ms{HTML Handling} corresponds to the requirement
\ms{R-51-HTML-HANDLING-OF-RDF-} \ms{LITERALS}.
Check if there are no HTML tags included in literals (of specific data properties within the context of specific classes).

@@

Check if literals of the property \ms{p} of the class \ms{C} do not include HTML tags.

@@

\ms{Negative Property Constraints} corresponds to the requirements
\begin{itemize}
	\item \ms{R-52-NEGATIVE-OBJECT-PROPERTY-CONSTRAINTS} and
	\item \ms{R-53-NEGATIVE-DATA-PROPERTY-CONSTRAINTS}.
\end{itemize}
Instances of a specific class must not have some object property.
In OWL 2 DL, this can be expressed as follows: \ms{ObjectComplementOf ( ObjectSomeValuesFrom ( ObjectPropertyExpression owl:Thing ) )}.

@@

\begin{ex}
# ShEx:
<FeelingForce> {
    feelingForce (true) ,
    attitute xsd:string }
<JediMentor> {
    feelingForce (true) ,
    attitute ('good') ,
    laserSwordColor xsd:string ,
    numberLaserSwords xsd:nonNegativeInteger ,
    mentorOf @<JediStudent> ,
   !studentOf @<JediMentor> }
<JediStudent> {
    feelingForce (true) ,
    attitute ('good') ,
    laserSwordColor xsd:string ,
    numberLaserSwords xsd:nonNegativeInteger ,
   !mentorOf @<JediStudent> ,
    studentOf @<JediMentor> }
\end{ex}

A matching triple has any predicate except those excluded by the '!' operator.

\begin{ex}
# individuals matching 'FeelingForce' and 'JediMentor' data shapes:
Obi-Wan 
    feelingForce true ;
    attitute 'good' ;
    laserSwordColor 'blue' ;
    numberLaserSwords '1'^^xsd:nonNegativeInteger ;
    mentorOf Anakin .
\end{ex}

\begin{ex}
# individuals matching 'FeelingForce' and 'JediStudent' data shapes:
Anakin 
    feelingForce true ;
    attitute 'good' ;
    laserSwordColor 'blue' ;
    numberLaserSwords '1'^^xsd:nonNegativeInteger ;
    studentOf Obi-Wan .
\end{ex}

@@

\ms{Handle RDF Collections} corresponds to the requirement
\ms{R-120-HANDLE-RDF-} \ms{COLLECTIONS}.
Examples are:
\begin{itemize}
	\item size of collection
	\item first / last element of list must be a specific literal
	\item compare elements of collection
	\item are collections identical?
	\item actions on RDF lists\footnote{See \url{http://www.snee.com/bobdc.blog/2014/04/rdf-lists-and-sparql.html}}
	\item 2. list element equals 'XXX'
	\item Does the list have more than 10 elements?
\end{itemize}

@@

Get 2. list element:

\begin{ex}
# SPIN:
getListItem
    a spin:Function ; rdfs:subClassOf spin:Functions ;
    spin:constraint [
        rdf:type spl:Argument ;
        spl:predicate sp:arg1 ;
        spl:valueType rdf:List ;
        rdfs:comment "list" ; ] ;
    spin:constraint [
        rdf:type spl:Argument ;
        spl:predicate sp:arg2 ;
        spl:valueType xsd:nonNegativeInteger ;
        rdfs:comment "item position (starting with 0)" ; ] ;
    spin:body [
        a sp:SELECT ;
        sp:text """
            SELECT ?item
            WHERE {
                ?arg1 contents/rdf:rest{?arg2}/rdf:first ?item } """ ; ] ;
    spin:returnType rdfs:Resource .
\end{ex}

\begin{ex}
# data:
Jinn students 
     ( Xanatos Kenobi ) . 
\end{ex}

\begin{ex}
# SPIN:
BIND ( getListItem( ?list, "1"xsd:nonNegativeInteger ) AS ?listItem ) .
\end{ex}

@@

\begin{ex}
# result:
Kenobi
\end{ex}

@@

\ms{Recursive Queries} corresponds to the requirement
\ms{R-222-RECURSIVE-QUERIES}.
If we want to define Resource Shapes, remember that it is a recursive
language (the valueShape of a Resource Shape is in turn another
Resource Shape). There is no way to express that in SPARQL without
hand-waving "and then you call the function again here" or "and then
you embed this operation here" text.  The embedding trick doesn't work
in the general case because SPARQL can't express recursive queries,
e.g. "test that this Issue is valid and all of the Issues that
references, recursively".
Most SPARQL engines already have
functions that go beyond the official SPARQL 1.1 spec. The cost of that
sounds manageable.

@@

\begin{ex}
# ShEx:
IssueShape {
    related @IssueShape*
}
\end{ex}

@@

@@

A date is really a date, as an example.
SPARQL regex can be used for this purpose.

@@

\begin{ex}
# SPIN:
FILTER ( datatype( ?shoeSize ) = xsd:nonNegativeInteger )
isNumeric ( ?shoeSize )
\end{ex}

The datatype of ?showSize is xsd:nonNegativeInteger.
The datatype is really numeric.

@@

%\subsection{Simple Example}
%
%\begin{DL} 
%$\rho(\textsf{hasBirthday}) \sqsubseteq \PVerb{xsd:dateTime}$\\
%\end{DL}
%
%Equal to the definition of a data property range.

@@

\begin{ex}
# OWL 2:
hasBrother owl:equivalentProperty hasMaleSibling . 
Chris hasBrother Stewie . 
Stewie hasMaleSibling Chris .
\end{ex}

@@

\begin{ex}
Chris hasMaleSibling Stewie . 
Stewie hasBrother Chris .
\end{ex}

\begin{DL}
hasBrother $\sqsubseteq$ hasMaleSibling $\sqcap$ hasMaleSibling $\sqsubseteq$ hasBrother \\
\end{DL}

@@

\ms{{Property Assertions}} corresponds to the requirement \ms{R-96-PROPERTY-ASSERTIONS}
and includes positive property assertions and negative property assertions.
A positive object property assertion ObjectPropertyAssertion( OPE a1 a2 ) states that the individual a1 is connected by the object property expression OPE to the individual a2. 
A negative object property assertion NegativeObjectPropertyAssertion( OPE a1 a2 ) states that the individual a1 is not connected by the object property expression OPE to the individual a2. 
A positive data property assertion DataPropertyAssertion( DPE a lt ) states that the individual a is connected by the data property expression DPE to the literal lt. 
A negative data property assertion NegativeDataPropertyAssertion( DPE a lt ) states that the individual a is not connected by the data property expression DPE to the literal lt.

@@

\begin{ex}
# OWL 2:
NegativeObjectPropertyAssertion( hasSon Peter Meg )
\end{ex}

@@

\begin{ex}
# OWL 2:
DataPropertyAssertion( :hasAge :Meg "17"^^xsd:integer )
\end{ex}

@@

\begin{ex}
# OWL 2:
hasFather rdf:type owl:FunctionalProperty . 	
Stewie hasFather Peter . 	
Stewie hasFather Peter_Griffin . 
\end{ex}

@@

\begin{ex}
Peter_Griffin owl:sameAs Peter .  
\end{ex}

@@

An object property inverse functionality axiom InverseFunctionalObjectProperty( OPE ) states that the object property expression OPE is inverse-functional - that is, for each individual x, there can be at most one individual y such that y is connected by OPE with x. Each such axiom can be seen as a syntactic shortcut for the following axiom: SubClassOf( owl:Thing ObjectMaxCardinality( 1 ObjectInverseOf( OPE ) ) ).

@@

\begin{ex}
# OWL 2:
fatherOf rdf:type owl:InverseFunctionalProperty . 	
Peter fatherOf Stewie .
Peter_Griffin fatherOf Stewie .
\end{ex}

@@

Entailment:

\begin{ex}
Peter owl:sameAs Peter_Griffin . 
\end{ex}

@@

\begin{ex}
# OWL 2:
Peter fatherOf Stewie . 
ObjectHasValue( fatherOf Stewie )
\end{ex}

The has-value class expression contains those individuals that are connected through the \ms{fatherOf} property with the individual \ms{Stewie}.
\ms{Peter} is classified as its instance.

@@

\begin{ex}
# OWL 2:
Peter likes Peter . 
ObjectHasSelf( likes ) 
\end{ex}

\begin{DL}
LikesThemselves $\sqsubseteq$ $\exists$ likes.Self
\end{DL}

@@

The self-restriction contains those individuals that like themselves.
\ms{Peter} is classified as its instance.

@@

\ms{Data Property Facets} corresponds to the requirement
\ms{R-46-CONSTRAINING-FACETS}.
For datatype properties it should be possible to declare frequently needed "facets" to drive user interfaces and validate input against simple conditions, including min/max value, regular expressions, string length etc. similar to XSD datatypes. 
Constraining facets to restrict datatypes of RDF literals.
Constraining facets may be: xsd:length, xsd:minLength, xsd:maxLength, xsd:pattern, xsd:enumeration, xsd:whiteSpace, xsd:maxInclusive, xsd:maxExclusive, xsd:minExclusive, xsd:minInclusive, xsd:totalDigits, xsd:fractionDigits.

@@

\begin{ex}
# OWL 2 QL (functional-style syntax):
Declaration( Datatype( SSN ) ) 
DatatypeDefinition( 
    SSN
    DatatypeRestriction( xsd:string xsd:pattern "[0-9]{3}-[0-9]{2}-[0-9]{4}" ) )     
DataPropertyRange( hasSSN SSN ) 
\end{ex}

\begin{ex}
# OWL 2 QL (turtle syntax):
SSN 
    a rdfs:Datatype ;
    owl:equivalentClass [
        a rdfs:Datatype ;
        owl:onDatatype xsd:string ;
        owl:withRestrictions ( 
            [ xsd:pattern "[0-9]{3}-[0-9]{2}-[0-9]{4}" ] ) ] .
hasSSN rdfs:range SSN .
\end{ex}

@@

%\subsection{Simple Example}
%Each object is uniquely identified by its social security number.
%
%\begin{ex}
%# OWL2 DL:
%owl:Thing owl:hasKey ( hasSSN ) . 	
%\end{ex}
%
%The first axiom makes hasSSN the key for instances of the owl:Thing class; thus, only one individual can have a particular value for hasSSN. 
%
%\begin{ex}
%# data
%Peter hasSSN "123-45-6789" . 
%PeterGriffin hasSSN "123-45-6789" . 
%\end{ex}
%
%\begin{DL}
%$(\textsf{funct } \textsf{SSN}\sp{\overline{\ }})$
%\end{DL}
%
%\begin{gcotable}
%property & $\top$ & SSN$^{-}$ & SSN & - & inverse & - \\
%property & $\top$ & SSN$^{-}$ & - & - & functional & - \\
%\end{gcotable}
%
%Since the values of hasSSN are the same for the individuals Peter and PeterGriffin, these two individuals are equal - that is, this ontology entails the following assertion Peter owl:sameAs PeterGriffin .

@@

%Since "Each object is uniquely identified by its social security number." It means it is inverse functional property.
%
%So instead, a property can be defined as 
%\begin{DL}
%$(\textsf{funct } \textsf{SSN})$
%\end{DL}
%, and then by a property inclusion axiom
%
%\begin{DL}
%$haskey \sqsubseteq SSN^{-}$
%\end{DL}
%
%now for a given number haskey will give me the object (note that obviously it should be named, no ambigious individuals).
%
%However I needed to do some little research and found out that, keys are even more general (A generalisation of inverse functional properties) so that it can also be a chain of properties. (it is clear here: http://www.w3.org/TR/2012/REC-owl2-rdf-based-semantics-20121211/\#table-semcond-keys)
%
%For this generalisation purposes and many spooky interactions of different sorts of key (and with some elaboration they can lead to undecidability), they introduced "key boxes, and extended the language with special "keyfor" construct. So how it is done in DL community as I learned is
%
%\begin{DL}
%SSN "keyfor" Human $\sqcap$ TurkishCitizen
%\end{DL}
%
 %for instance. Now ssn behaves as an identifying element for individuals of right hand side.
%
%For comprehensive studies and resources, one can find lots of examples in:
%
%http://lat.inf.tu-dresden.de/research/mas/Din-Mas-06.pdf
%https://www.jair.org/media/1542/live-1542-2375-jair.pdf
%
%Yes you can, but as you see it is not restriction statement by its own. Nevertheless, I think, in DL mapping, use of "keyfor" and sending a reference to one of those papers is more convenient/safer, since it can be a datatype property, object property, or a chain of properties (from OWL world), as well as DL literature has it as "keyfor". 
%
%\begin{gcotable}
%property & $\top$ & SSN & - & - & keyfor & - \\
%\end{gcotable}

The \ms{Primary Key Properties} constraint is often useful to declare a given (datatype) property as the "primary key" of a class, so that a system can enforce uniqueness. 
Starfleet officers, e.g., are uniquely identified by their command authorization code (e.g. to activate and cancel auto-destruct sequences).
It means that the property \ms{commandAuthorizationCode} is inverse functional - mapped to DL and the RDF-CO as follows:

\begin{DL}
$(\ms{funct } commandAuthorizationCode\sp{\overline{\ }})$
\end{DL}

\begin{gcotable}
property & $\top$ & commandAuthorizationCode$^{-}$ & commandAuthorizationCode$^{-}$ & - & inverse & - \\
property & $\top$ & commandAuthorizationCode$^{-}$ & - & - & functional & - \\
\end{gcotable}

Keys, however, are even more general, i.e., a generalization of inverse functional properties \cite{Schneider2009}.
A key can be a datatype property, an object property, or a chain of properties.
For this generalization purposes, as there are different sorts of key, and as keys can lead to undecidability, 
DL is extended with \ms{key boxes} and a special \ms{keyfor} construct\cite{Lutz2005}.
This leads to the following DL and RDF-CO mappings (only one simple property constraint):

\begin{DL}
commandAuthorizationCode \ms{keyfor} StarfleetOfficer
\end{DL}

\begin{gcotable}
property & StarFleetOfficer & commandAuthorizationCode & - & - & keyfor & - \\
\end{gcotable}

@@

\ms{Exploiting Class/Property Specialization Ontology Axioms} corresponds to the requirement
\ms{R-224-instance-level-data-validation-exploitingclass/} \ms{property-specialization-axioms-in-ontologies}.
Validation of instances data (direct or indirect) exploiting the subclass or sub-property link in a given ontology.
This validation can indicate when the data is verbose (redundant) or expressed at a too general level, and could be improved.
Examples are:
\begin{itemize}
	\item If dc:date and one of its sub-properties dcterms:created or dcterms:issued are present, check that the value in dc:date is not redundant with dcterms:created or dcterms:issued for ingestion
  \item Check if dc:rights has the same value than edm:rights either as rdf:resource or literal, if yes dc:rights is redundant
  \item If one or more dc:coverage are present, suggest the use of one of its sub-properties, dcterms spatial or dcterms:temporal.
\end{itemize}
